---
title: "Initial Korean covid 19"
author: "Jayline Zelaya"
date: "todays date"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
```{r}
structure(kRdata)
```


  
# Your Question - Is dying independent of sex?

$H_0$ Dying is not independent of sex

$H_A$ Dying is independent of sex

## METHOD

Sex is categorical data and result is categorical data, therefore we use cat~cat helper file.



## Graphical results - use ggplot commands to do this 

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=Sex,y=count,fill=result))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=Sex,y=prcnt*100,fill=result)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This bar chart shows of females 55% who were infected with covid are still alive and over 35% are dead. Of males who were infected with covid slightly over 40% are alive and over 60% are dead.
  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

This table shows the percentage of females alive is 98.7 and males is 97% of those who live and 1.4% of females infected died and 3% of males infected died .

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Chisq table shows the P-value is 0.001 Fisher exact test says P value is .0009 therefore we reject the null hypothesis.

## Conclusions

In conclusion there are more females who stayed alive than males. and dying is independent of sex.
